---
title: "generating neighborhood divisions markdown"
author: "kira"
date: "11/22/2021"
output:
  html_document:
     toc: true
     theme: cerulean
     highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      fig.width=6, fig.height=5
                      ,message = FALSE
                      ,warning = FALSE)

proj.dir <- #"~/R/all sharkey geoseg work/divseg/R/visuals & mkdwns/"
  "~/R/all sharkey geoseg work/divM/"
# ---------------------------------------------------------------------------
rm(list = ls())
require(tidyverse)
require(sf)
library(mapview)

# option setting
sf_use_s2(T)
options(tigris_use_cache = TRUE)

devtools::load_all()
```

## Ascribing Neighborhood dividedness

This markdown shows the workflow and functions for allocating neighborhood (nbhd) dividedness. 

This package bundles a couple functions to query nbhds within a region, and allocate each to a sub-area bounded by a type of division. For example, localities or school districts within a metro area, sides of a interstate/highway relative to one another, or other linear divisions that might be considered.


### Sample call

#### CZ

The function can be called like:

```{r sample}

# I use my helper library geox (see below for installation)

# define a place to get nbhd divisions for 
czsf <- geox::build.CZs('20100') # Portland, ME commuting zone

# get divisions that can be overlaid 
hwys <- geox::get.NHPN(sfx = st_transform(czsf, 4326))

devtools::load_all()

# allocate census tracts to sub-area divisions, using just interstates
nb.div <- divM::gen.cross.tract.dividedness(
  region = czsf
  ,divs = filter(hwys, signt1 == 'I')
  ,nbd.query.fcn = tigris::tracts
  ,fill.nhpn.gaps = T
  ,region.id.colm = 'cz'
  ,erase.water = F
  ,year = 2019
  )

# this object represents neighborhood dividedness in the region.
nb.div

```

The `nb.div` object, which represents neighborhood dividedness in the region has a `geoid` column for tract or block group identifier, a `poly.id` column for the sub-polygon division, and a `perc.in.div` column for the percentage of the neighborhood within the identified sub-polygon division. 

We can visualize the neighborhood allocation like:

```{r smplczvis}

nb.divsf <- nb.div %>% geox::attach.geos(query.fcn = tigris::tracts, year = 2019)

ggplot() +
  geom_sf(data = nb.divsf
          ,aes(fill = poly.id)
          ,color = 'white', size = .6
          ,alpha = .5) +
  geom_sf(data =  filter(hwys, signt1 == 'I')
          ,aes(color = signt1)
          ,size = 1.3) +
  #scale_fill_discrete(guide= 'none') +
  theme_void()

```

The pre-generated cross-tract or cross-blockgroup dividedness datasets are generated using Della scripts that just iterate through different regions.

The example, and the function itself, lean on `geox`, a helper library I wrote that has a lot of convenience functions working with census spatial data, querying through the `tigris` library, and referencing CZs and CBSAs.

To install this library, use:
`devtools::install_github("https://github.com/kmcd39/geox.git")`
Github link: https://github.com/kmcd39/geox/tree/main/R

#### City-level variation

Variations are easy to generate, with different regions or divisions objects. For example, here's a separate generation block-group dividedness within the city of Portland (rather than tracts within the commuting zone). We can also use all highways instead of just interstates

```{r plc.sample}
# get all places in the Portland CZ
plcs <- geox::places.wrapper(x = czsf)

# filter to just portland
smplc <- plcs %>%
  filter(grepl('^Portland', name)) %>%
  st_transform(4326)

# allocate block groups to sub-area divisions, using all hwys
nb.plc.div <- divM::gen.cross.tract.dividedness(
  region = smplc
  ,divs = hwys
  ,nbd.query.fcn = tigris::block_groups
  ,fill.nhpn.gaps = T
  ,region.id.colm = 'placefp'
  ,erase.water = F
  ,year = 2019
  )

nb.plc.div

# get background tiles for vis this time
sttm <- visaux::get.stamen.bkg(smplc, zoom = 12)

nb.plc.divsf <- nb.plc.div %>% geox::attach.geos(query.fcn = tigris::block_groups)

ggmap(sttm) +
  geom_sf(
    data = nb.plc.divsf
    ,aes(fill = poly.id)
    ,color = 'white'
    ,alpha = .6
    ,inherit.aes = F
  ) +
  geom_sf(data = hwys
          ,aes(color = factor(signt1))
          ,size =  1.3
          ,inherit.aes = F
  ) +
  scale_fill_discrete(guide = 'none') +
  visaux::bbox2ggcrop(nb.plc.divsf) +
  theme_void() +
  theme(legend.position = 'bottom')


```

See the R script "sample polydiv gen.R" to see a longer walkthrough of the workflow that generates this.
